---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First, we unzip the data that is stored in the "activity.zip" file and load it to R.

```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv", header = TRUE)
```
And we check if it is necesary to process the data. First we check summary and structure of the data:

```{r}
summary(activity)
str(activity)
```
We make interval column into a factor so we can use it later to make the plots

```{r}
activity$interval <- as.factor(activity$interval)
```

And convert the date column into a date format

```{r}
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```
## What is mean total number of steps taken per day?

To answer this question, we first compute the total steps taken per day

```{r}
stepsxd <- sapply(split(activity, activity$date), function(x){ sum(x[,1], na.rm = TRUE)})
```
We make a data frame so we can plot it

```{r}
df <- data.frame(stepsxd, days = unique(as.character(activity$date)))
```

and we make a histogram with this data 

```{r}
barplot(stepsxd~days, data=df, ylab = "Total of steps", main="Total of steps per day")
```

And now we compute the mean and median of steps per weekday

```{r}
df$days <- as.Date(df$days, format = "%Y-%m-%d")
meansteps <- sapply(split(df, weekdays(df$days)), function(x){mean(x[,1])})
mediansteps <- sapply(split(df, weekdays(df$days)), function(x){median(x[,1])})
data.frame(mean=meansteps, median=mediansteps)
```


## What is the average daily activity pattern?

First we compute the average of steps taken acrooss the 5 minute intervals

```{r}
avgsteps <- sapply(split(activity, activity$interval), function(x){mean(x[,1], na.rm = TRUE)})
df2 <- data.frame(interval=1:288, avgsteps)
```
and then we make the time series plot

```{r}
plot(df2, ylab="average of steps", type="l")
activity$interval[100]
```

In this plot we can see that the 5-minute interval in which more steps are taken is around the interval 100, that is in the interval between 8:15 and 8:20.


## Imputing missing values

The missing values in the data sets are:
```{r}
sum(is.na(activity$steps))
```
We can see that there are several days with no count of steps, and we notice this since very day must contain 288 observations.

```{r}
sapply(split(activity, activity$date), function(x){sum(is.na(x[,1]))})
```

In order to fill in the missing values, we first duplicate the data frame and we replace the mssing values with -1's and we find the indices where the activity dataset has missing values.

```{r}
activity_duplicate <- activity
activity_duplicate[is.na(activity)] <- -1
na <- is.na(activity$steps)
```

Next, we make into 0 the indices where activity has no missing values

```{r}
activity_duplicate$steps[!na] <- 0
```

Now, we introduce the averaged steps per 5-minute interval in the new dataset.

```{r}
activity_duplicate$steps2 <- activity_duplicate$steps*avgsteps
activity_duplicate$steps2 <- (-1)*activity_duplicate$steps2
activity_duplicate$steps <- activity$steps
activity_duplicate$steps[na] <- 0
```

Next, we add original data and new data

```{r}
activity_duplicate$steps2 <- activity_duplicate$steps+activity_duplicate$steps2
```

And finally we obtain our new dataset, that has no missing values

```{r}
activity_2 <- data.frame(steps=activity_duplicate$steps2, activity[,2:3])
sum(is.na(activity_2$steps))
```

Now, we repeat the process to make a new plot:

```{r}
stepsxd <- sapply(split(activity_2, activity$date), function(x){ sum(x[,1])})
df <- data.frame(stepsxd, days = unique(as.character(activity_2$date)))
barplot(stepsxd~days, data=df, ylab = "Total of steps", main="Total of steps per day")
```

To calculate and report mean and median of the new data, we repeat the previous process


## Are there differences in activity patterns between weekdays and weekends?
